### 2013-12-30 conclusions.tex
fixed some quantitative data

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations between estimated organ
volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
$\pm$ \ARFImrTotalLatLatStdPct\%), while over-estimates of the other dimensions
were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
bias between the imaging modalities, but significant variability case-to-case
(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
volume differences were, again, strongly attributed to over-estimation of the
lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
anterior-to-posterior dimension (\ARFImrCentralAntPostMeanPct~$\pm$
\ARFImrCentralAntPostStdPct\%).  Strong variability in central gland volumes is
believed to be related to the extent of benign prostatic hyperplasia (BPH) for
select cases.  Overall, ARFI imaging of the prostate yielded prostate volumes
and dimensions that were correlated with MR T2WI estimates, with biases in the
anterior-to-posterior dimension, most likely related to poor displacement SNR
in the anterior region of the prostate from greater distance from the rectal
wall imaging surface, and the lateral dimension, where contrast between the PZ
and peri-prostatic fat could be limited.  ARFI imaging is a promising low-cost,
real-time imaging modality that can compliment MR imaging for diagnosis,
treatment planning and management of PCa.

>>>>

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations between estimated organ
volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
$\pm$ \ARFImrTotalLatLatStdPct\%), while differences in the other dimensions
were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
bias between the imaging modalities, but significant variability case-to-case
(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
volume differences were, again, strongly attributed to over-estimation of the
lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
apex-to-base dimension (\ARFImrCentralAntPostMeanPct~$\pm$
\ARFImrCentralAntPostStdPct\%).  Strong variability in CG volumes is believed
to be related to the extent of BPH for select cases.  Overall, ARFI imaging of
the prostate yielded prostate volumes and dimensions that were correlated with
MR T2WI estimates, with differences in the lateral and apex-to-base dimensions.
Poor image contrast between the prostate and peri-prostatic fat may be one
cause of these differences.  ARFI imaging is a promising low-cost, real-time
imaging modality that can compliment MR imaging for diagnosis, treatment
planning and management of PCa.

### 2013-12-30 discussion.tex
fixed some quantitative data

\section{Discussion}\label{sect:discussion}
This study was designed to compared the zonal anatomy of the prostate in
patients with biopsy-proven PCa as seen in ARFI imaging and MR T2WI.  Overall,
ARFI imaging performed well when compared with MR estimates of total and
central gland volume (Figure~\ref{fig:mr_arfi_volumes}), though ARFI tended to
overestimate prostate total gland volume, most commonly due to over-estimation
of the lateral dimension of the prostate.  This can be due to poor contrast
between the PZ and peri-prostate fat (Figure~\ref{fig:arfi_segs}).  As seen in
Figure~\ref{fig:arfi_segs}, clear delineation of the anterior aspect of the
prostate can also be challenging in ARFI and B-mode images due to poor SNR,
since this is the deepest aspect of the prostate relative to the rectal wall
imaging surface. MR T2WI can also have difficulties visualizing and diagnosing
lesions in the anterior region of the prostate,~\cite{Gupta2013}, which is
thought to be secondary to the heterogeneity of the gland in that area,
especially in the presence of BPH.  This was evident in imaging study subject
4, with a $\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}), which was an
outlier in terms of weight and volume.  The most accurate measurements for both
the total prostate gland and central gland were in the apex-to-base dimension
(Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}),
with no mean bias in ARFI imaging estimates, but significant case-to-case
variability, leading to poor linear regression coefficients of determination.

This study has several limitations that should be considered when interpreting
these results.  Gross pathology weight and axis measurements could both be
affected by the presence of periprostatic tissue that was excised during
radical prostatectomy, especially in cases where more aggressive margins may
have been necessary.  For this reason, unlike the image-to-image measurement
comparisons, all of the image metrics presented relative to pathology metrics
(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
but instead, relative correlations were evaluated.  Additionally, the volumes
of the prostate from gross pathologic measurements were approximated as
ellipsoids, which also introduced error, most likely an over-estimation of
volume.  Interestingly, all pathology estimates were thought to have positive
biases, but both imaging modalities tended to overestimate relative to the
pathology measurements.

It should also be noted that all of the prostates in this study contained
varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
anatomy, especially in the case of BPH and central gland morphology.  While
younger, healthier prostates could have been targeted, these healthy organs
would not have been excised for pathology characterization, and the zonal
anatomy of a health (young) prostate is expected to be different from the
prostate of a middle-age man, who is the target demographic for PCa screening
imaging and PCa characterization; therefore, it was felt that performing this
analysis in the presence of these confounding factors was appropriate.

>>>>

\section{Discussion}\label{sect:discussion}
This study was designed to compared the zonal anatomy of the prostate in
patients with biopsy-proven PCa as seen in ARFI imaging and MR T2WI.  Overall,
ARFI imaging performed well when compared with MR estimates of total and
central gland volume (Figure~\ref{fig:mr_arfi_volumes}), though ARFI tended to
overestimate prostate total gland volume, most commonly due to over-estimation
of the lateral dimension of the prostate.  This can be due to poor contrast
between the PZ and peri-prostatic fat (Figure~\ref{fig:arfi_segs}).
Figure~\ref{fig:arfi_segs} also demonstrates challenges in delineating the
anterior aspect of the prostate in both ARFI and B-mode images due to poor SNR
since this is the deepest aspect of the prostate relative to the rectal wall
imaging surface. MR T2WI can also have difficulties visualizing and diagnosing
lesions in the anterior region of the prostate,~\cite{Gupta2013}, which is
thought to be secondary to the heterogeneity of the gland in that area,
especially in the presence of BPH.  This was evident in imaging study subject
4, with a $\sim$143 g prostate (Figure~\ref{fig:mr_arfi_weight}), which was an
outlier in terms of weight and volume.  While the anterior aspect of the
prostate can be challenging to visualize, it was the axis with the best
agreement between ARFI and MR imaging (Figure~\ref{fig:mr_arfi_path_axes}(b),
Table~\ref{tab:mr_arfi_axes_error}).  ARFI images consistently underestimated
the extent of the prostate capsule and CG along the apex-to-base axis relative
to MR T2WI (Figure~\ref{fig:mr_arfi_path_axes}(c),
Table~\ref{tab:mr_arfi_axes_error}), though the reasons for this
underestimation were not clear.

This study has several limitations that should be considered when interpreting
these results.  Gross pathology weight and axis measurements could both be
affected by the presence of periprostatic tissue that was excised during
radical prostatectomy, especially in cases where more aggressive margins may
have been necessary.  For this reason, unlike the image-to-image measurement
comparisons, all of the image metrics presented relative to pathology metrics
(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
but instead, relative correlations were evaluated.  Additionally, the volumes
of the prostate from gross pathologic measurements were approximated as
ellipsoids, which also introduced error, most likely an over-estimation of
volume.  Interestingly, all pathology estimates were thought to have positive
biases, but both imaging modalities tended to overestimate volume relative to the
pathology measurements.

It should also be noted that all of the prostates in this study contained
varying amounts of PCa, BPH and atrophy, all of which can distort the zonal
anatomy, especially in the case of BPH and central gland morphology.  While
younger, healthier prostates could have been targeted, these healthy organs
would not have been excised for pathology characterization, and the zonal
anatomy of a healthy (young) prostate is expected to be different from the
prostate of a middle-age man, who is the target demographic for PCa screening
imaging and PCa characterization.

### 2013-12-30 fig_mr_arfi_weight.tex
fixed some quantitative data

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostated weights (a,
    R$^2$ = 0.68).  T2WI MR (blue, X) showed a moderate correlation between the
    reconstructed volumes and prostate weight (R$^2$ = 0.44), while volumes
    reconstructed from ARFI images (green, O) showed weaker correlation (R$^2$
    = 0.21) (b).  Even weaker correlations existed between both T2WI MR and
    ARFI image volumens and approaximated ellipsoidal prostate pathology
    volumes (R$^2$ = 0.08 and 0.01, respectively) (c).  It should be noted in
    these figures that two study subjects had excessively large prostates that
    were difficult to fully capture in imaging (volumes $>$ 60 cm$^3$),
    especially in their anterior region, and they were, therefore, grossly
    underestimated in size.}
\label{fig:mr_arfi_weight}
\end{figure}

>>>>

\begin{figure}[htb!]
\centering
\begin{tabular}{lll}
\includegraphics[width=0.3\linewidth]{figs/corr_path_vol_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol} &
\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol} \\
(a) Path Weight : Path Volume & (b) Image Volume : Prostate Weight & (c) Image Volume : Path Volume \\
\end{tabular}
%\begin{tabular}{ll}
%\includegraphics[width=0.3\linewidth]{figs/corr_weight_vol_no4} &
%\includegraphics[width=0.3\linewidth]{figs/corr_pathVol_vol_no4} \\
%(d) Image Volume : Prostate Weight (-4) & (e) Image Volume : Path Volume (-4) \\
%\end{tabular}
\caption{Tri-axial pathology measurements were used to make an ellipsoidal
    prostate volume approximation based on gross pathology axis measurements,
    which was moderately well-correlated with the excised prostated weights (a,
    R$^2$ = \pathVolWeightRsq).  T2WI MR (blue, X) showed a moderate
    correlation between the reconstructed volumes and prostate weight (R$^2$ =
    \weightMRrsq), while volumes reconstructed from ARFI images (green, O)
    showed weaker correlation (R$^2$ = \weightARFIrsq) (b).  Even weaker
    correlations existed between both T2WI MR and ARFI image volumens and
    approaximated ellipsoidal prostate pathology volumes (R$^2$ = \pathVolMRrsq
    and \pathVolARFIrsq, respectively) (c).  It should be noted in these
    figures that two study subjects had excessively large prostates that were
    difficult to fully capture in imaging (volumes $>$ 60 cm$^3$), especially
    in their anterior region, and they were, therefore, grossly underestimated
    in size.}
\label{fig:mr_arfi_weight}
\end{figure}

### 2013-12-30 image_proc_methods.tex
fixed some quantitative data

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base to apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG and urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogenous high signal intensity on T2WI, which is usually
similar to that of the nearby periprostatic fat. The CG was visualized and
delineated based on its heterogeneous and lower signal intensity as well as its
location (Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG
segmentation. Although not readily visible on every case, the AFS was
identified by its low T2 signal intensity and its location anterior to the
central gland.

@@

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
using the \verb+draweffect+ tool to model and measure the prostate capsule and CG
volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
segmentation because it has a greater number of easily identifiable anatomical
structures:  seminal vesicles in the left and right posterior base, urethra in
the mid apex, surrounding fibromuscular tissue in the anterior , and rectum in
the posterior  than either the axial or coronal views. Prostate capsule was
segmented in B-mode as this modality offers both high contrast between
prostatic tissue and surrounding periprostatic fat, and better SNR in the
anterior capsule region than ARFI images. ARFI was used to segment central gland as
there is clear delineation in this modality between central gland and
peripheral zone. This delineation is a result of the greater density of
connective tissue within the central gland versus the peripheral zone meaning
the central gland is stiffer than the surrounding peripheral zone.  Prior to
segmentation, both the B-mode and ARFI image stacks were downsampled along the
sagittal axis by a factor of 20 (i.e., from 360 slices per image stack to 18
slices per image stack). This was done because segmenting 360 slices per image
is inordinately time intensive, and does not improve the quality of the 3D
rendered model any more than segmenting the more reasonable 18 slices per image
stack. The parameters used for the resample scalar volume tool are shown below:

WHAT GOES HERE?

Although, segmentations were performed along the sagittal plane, label map
intersections in the axial and coronal views were used during active creation
of the sagittal segmentation to further improve segmentation accuracy.  This
was done by using the resampled image stack, B-mode for capsule and ARFI for
CG, as the background in the sagittal plane, and the original image stack as
background in the axial and coronal planes. All three orthogonal views were set
to the label map assigned to the resampled volume. The B-mode derived capsule
was segmented and modeled before the central gland. Parameters used for 3D
model rendering were identical to those used in table X for MR modeling. The
modeled capsule  helprf guide the CG segmentation by enabling capsule model
intersections in the three orthogonal planes during CG segmentation. The
capsule provided an outer boundary for the CG segmentation. The capsule and CG
models were then measured identically to the MR measurement procedure.

>>>>

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented using the smooth polygon tool in
ITK-SNAP~\cite{Yushkevich2006}, using unique labels for the PZ, CG and AFS. The
gland was segmented from base to apex.  The base was identified below the
bladder, and subsequent images were segmented until the last slice with visible
prostatic tissue was identified caudally. The CG, urethra, PZ and AFS were
segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby periprostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.

@@

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
using the \verb+draweffect+ tool to model and measure the prostate capsule and
CG volume and tri-axial dimensions.  The sagittal imaging plane was chosen for
segmentation because it has a greater number of easily identifiable anatomical
structures:  seminal vesicles in the left and right posterior base, urethra in
the mid apex, surrounding fibromuscular tissue in the anterior , and rectum in
the posterior  than either the axial or coronal views. Prostate capsule was
segmented in B-mode as this modality offers both high contrast between
prostatic tissue and surrounding periprostatic fat, and better SNR in the
anterior capsule region than ARFI images. ARFI images were used for CG
segmentation.  Prior to segmentation, both the B-mode and ARFI image stacks
were down-sampled along the sagittal axis by a factor of 20 (i.e., from 360
slices per image stack to 18 slices per image stack) to reduce the number of
manually-segmented frames.

While segmentations were performed along the sagittal plane, label map
intersections in the axial and coronal views were used during active creation
of the sagittal segmentation to further improve segmentation accuracy.
Parameters used for 3D model rendering were identical to those used for MR
modeling (Table~\ref{tab:3dslicer}).  The modeled capsule helped guide the CG
segmentation by enabling capsule model intersections in the three orthogonal
planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule provided
an outer boundary for the CG segmentation. The capsule and CG models were then
measured using the same procedure as the MR images.

### 2013-12-30 results.tex
fixed some quantitative data

Measurements of the prostate total and CG dimensions along the three standard
anatomic axes (apex-to-base, lateral-to-lateral, and anterior-to-posterior)
were made (Table~\ref{tab:mr_arfi_axes}), and the correlation between the
imaging axis measurements was analyzed (Figure~\ref{fig:mr_arfi_path_axes}).
ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
and CG (R$^2$ = \totalLatLatRsq~and \centralLatLatRsq, respectively), with mean
overestimates of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
= \totalAntPostRsq~and \centralAntPostRsq), respectively, with overestimates of
\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct and
\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, central gland),
with differences of \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct~and \ARFImrCentralApexBaseMeanPct~$\pm$
\ARFImrCentralApexBaseStdPct\%, respectively.

>>>>

Measurements of the prostate total and CG dimensions along the three standard
anatomic axes (apex-to-base, lateral-to-lateral, and anterior-to-posterior)
were made (Table~\ref{tab:mr_arfi_axes}), and the correlation between the
imaging axis measurements was analyzed (Figure~\ref{fig:mr_arfi_path_axes}).
ARFI was most correlated to MR in the lateral-to-lateral axis in both the total
and CG (R$^2$ = \totalLatLatRsq~and \centralLatLatRsq, respectively), with mean
overestimates of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
= \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG),
with underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$
\ARFImrCentralApexBaseStdPct\%, respectively.
